---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FastKRR

<!-- badges: start -->
<!-- badges: end -->

The goal of FastKRR is to fit a Kernel Ridge regression estimator based on a set of data $(x_i, y_i)^n_{i = 1}$.

**Dependencies:** Rcpp, RcppArmadillo, CVST, parsnip  
This package uses **CVST** (GPL ≥ 2). Overall license: **GPL (≥ 3)**.

## Installation

You can install the development version of FastKRR from [GitHub](https://github.com/kybak90/FastKRR) with:

``` r
# install.packages("pak")
pak::pak("kybak90/FastKRR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(FastKRR)

# example data set
set.seed(1)
n = 1000; d = 1
rho = 1
X = matrix(runif(n*d, 0, 1), nrow = n, ncol = d)
y = as.vector(sin(2*pi*rowMeans(X)^3) + rnorm(n, 0, 0.1))

# model fitting - exact
model_exact = fastkrr(X, y, kernel = "gaussian", rho = rho, opt = "exact")

# model fitting - pivoted
model_pivoted = fastkrr(X, y, kernel = "gaussian", rho = rho, opt = "pivoted")

# model fitting - nystrom
model_nystrom = fastkrr(X, y, kernel = "gaussian", rho = rho, opt = "nystrom")

# model fitting - rff
model_rff = fastkrr(X, y, kernel = "gaussian", rho = rho, opt = "rff")


# predict
new_n = 500
new_x = matrix(runif(new_n*d, 0, 1), nrow = new_n, ncol = d)
new_y = as.vector(sin(2*pi*rowMeans(new_x)^3) + rnorm(new_n, 0, 0.1))

pred_exact = pred_krr(model_exact, new_x)
pred_pivoted = pred_krr(model_pivoted, new_x)
pred_nystrom = pred_krr(model_nystrom, new_x)
pred_rff = pred_krr(model_rff, new_x)
```


You can also embed plots, for example:

```{r echo = FALSE}
library(ggplot2)
data = data.frame(new_x, new_y)
data$pred_exact = as.numeric(pred_exact)
data$pred_pivoted = as.numeric(pred_pivoted)
data$pred_nystrom = as.numeric(pred_nystrom)
data$pred_rff = as.numeric(pred_rff)

ggplot(data = data, aes("x" = new_x, "y" = new_y)) +
  geom_point("x" = new_x, "y" = new_y) +
  geom_line(aes("x" = new_x, "y" = pred_exact, color = 'opt = "exact"'), linewidth=1.2) + 
  geom_line(aes("x" = new_x, "y" = pred_pivoted, color = 'opt = "pivoted"'),linewidth=1.2) + 
  geom_line(aes("x" = new_x, "y" = pred_nystrom, color = 'opt = "nystrom"'), linewidth=1.2) + 
  geom_line(aes("x" = new_x, "y" = pred_rff, color = 'opt = "rff"'), linewidth=1.2)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
